---
title: "Exercise prediction"
author: "VBolwerk"
date: "10 juni 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise prediction

The goal of this project is to predict the manner in which 6 participants did their exercises. This is the "classe" variable in the training set. We will use the other variables to predict with. In this report I will describe how I built my model, how I used cross validation, what I think the expected out of sample error is, and why I made the choices I did. I will also use the prediction model to predict 20 different test cases.

We start by loading in the data.

```{r}
library(randomForest)
library(caret)
library(AppliedPredictiveModeling)
train_df <-read.csv("~/../Desktop/pml-training.csv")
train_df$classe = as.factor(train_df$classe)

train_df <-train_df[, 7:160]
```

## Model selection and validation

First we make a cross validation splot, using a 20% validation set.


```{r}
nearzero = nearZeroVar(train_df, saveMetrics = TRUE)
cols_to_keep = row.names(nearzero[nearzero$nzv==FALSE,])
train_df = train_df[, cols_to_keep]

non_missing <- (colSums(is.na(train_df))==0)

train_df = train_df[, non_missing]


partition = createDataPartition(train_df$classe, p=0.8, list=FALSE)
training = train_df[partition, ]
validation = train_df[-partition, ]
```

Then we create a model, using random forest and 3 fold cross-validation. 

```{r}
model =train(classe ~ ., data=training, method="rf",trControl = trainControl(method = "cv", number=3))
```

We evaluate our model on the validation set below.

```{r}
validation_vars <- validation[, colnames(train_df)[1:53]]
predictions <-predict(model, validation_vars)
confusionMatrix(validation$classe, predictions)
```

We can see that our model works really well, so we are ready to go towards the test set. Out of sample accuracy was 99,9%, which means that the expected accuracy on our test set should also be high.

```{r}
test_df <-read.csv("~/../Desktop/pml-testing.csv")

test_df <-test_df[, colnames(train_df)[1:53]]


predictions <-predict(model, test_df)
predictions
```